---
title: "Covid-19 Cases Analysis in Toronto" 
author: "Jingyi Chen" 
date: "2020-09-27" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages(c('repr', 'IRdisplay', 'evaluate', 'crayon', 'pbdZMQ', 'devtools', 'uuid', 'digest')
#install.packages("visdat")
#install.packages("magrittr")
#devtools::install_github("sharlagelfand/opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(dplyr)
library(visdat)
library(magrittr)

package <- show_package("64b54586-6180-4485-83eb-81e8fae3b8fe")
package

# get all resources for this package
resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
data

```

# Abstract
In this article, we investigated the covid-19 cases in Toronto. We want to demonstrate how covid-19 cases spread around Toronto, what age group has the highest number of cases, and the rate of severity in all cases, etc. Through analyzing the data from open data Toronto, we want to show people a comprehensive and detailed circumstances of this ongoing pandemic in Toronto.


# Introduction

```{r, echo=FALSE, , include=FALSE}
visdat::vis_dat(data)

skimr::skim(data)

class(data$`Age Group`)

active_data <-
  data %>% 
  filter(Outcome == "ACTIVE")

fatal_data <-
  data %>% 
  filter(Outcome == "FATAL")

select_data <-
  data %>% 
  select(`Age Group`, `Outbreak Associated`, `Source of Infection`, `Client Gender`, Outcome)

```

## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}

data %>%
  ggplot(aes(x = `Age Group`)) +
  geom_bar() +
  labs(x = "Age",
       y = "Number",
       title = "Distribution of Age Goups in Covid-19 Cases.",
       tag = "Graph 1")+
   theme(axis.text.x = element_text(size = 4.5))    

data %>%
  ggplot(aes(x = `Client Gender`)) +
  geom_bar() +
  labs(x = "Gender",
       y = "Number",
       title = "Distribution of Gender Goups in Covid-19 Cases.")+
   theme_bw()
  
data %>%
  ggplot(aes(x = Outcome)) +
  geom_bar() +
  labs(x = "Outcome",
       y = "Number",
       title = "Distribution of Outcomes in Covid-19 Cases.")+
   theme_bw()

data %>%
  ggplot(aes(x = `Source of Infection`)) +
  geom_bar() +
  labs(x = "Source of Infection",
       y = "Number",
       title = "Distribution of Source of Infection in Covid-19 Cases.")+
   theme(axis.text.x = element_text(size = 4.5))


```

# Discussion
Most cases in which area
Most outbreak associated
Most patient in which age group
Most patients were infected by
More female patients than male patients
age group in each gender
Total fatal cases and active cases, most active cases in each area

How many patients are currently hospitalized
how many patients are ever hospitalized
How many patients are currently in ICU
How many patients are ever in ICU
How many patients are currently intubated
How many patients are ever intubated

percentage of fatality for patient ever in ICU, intubated and hospitalized.

correlation between age and conditions
```{r, echo=FALSE, warning = FALSE, message=FALSE}

  data %>%
    group_by(FSA) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
   
  data %>%
    group_by(`Neighbourhood Name`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
  
  data %>%
    group_by(`Age Group`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))

  data %>%
    group_by(`Source of Infection`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
  
  data %>%
    group_by(`Client Gender`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
  
  data %>%
    group_by(`Client Gender`, `Age Group`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
  
  data %>%
    group_by(Outcome) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
  
  data %>%
    filter(Outcome == "ACTIVE") %>%
    group_by(`Neighbourhood Name`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
    
  data %>%
    group_by(`Currently Hospitalized`) %>%
    summarise(count=n()) %>%
    arrange(desc(count))
```
## Weakness and next steps

# References
Carrieri, D., Peccatori, F. A., & Boniolo, G. (2020). COVID-19: a plea to protect the older population. International journal for equity in health, 19(1), 72. https://doi.org/10.1186/s12939-020-01193-5

D'cruz, M., & Banerjee, D. (2020). 'An invisible human rights crisis': The marginalization of older adults during the COVID-19 pandemic - An advocacy review. Psychiatry research, 292, 113369. Advance online publication. https://doi.org/10.1016/j.psychres.2020.113369


Jalili, M., Payandemehr, P., Saghaei, A., Sari, H. N., Safikhani, H., & Kolivand, P. (2020). Characteristics and Mortality of Hospitalized Patients With COVID-19 in Iran: A National Retrospective Cohort Study. Annals of internal medicine, M20-2911. Advance online publication. https://doi.org/10.7326/M20-2911
